Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

### 2022-09-20 paper/paper.Rmd
update with comments from @sbfnk

---
title: Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic. However, these forecasts require a coordinated team to produce and are effectively black boxes. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. Here we use observations of the behaviour of the ECDC forecast ensemble, combined with our own forecasting experience,  to specify a series of model assumptions. We then develop a forecast model using these assumptions and compare its performance to the ECDC forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon. We found that this surrogate model performed visibly similarly to a subset of example locations though with increased uncertainty. Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse but that median performance was more comparable. The majority of the difference in performance was linked to a few forecasts from a period of high incidence. We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict. The central part of the surrogate models forecast distribution performed more similarly to ensemble forecasts distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance. We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble but that more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements. We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
bibliography: library.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
---

@@

# Introduction {-}

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade with significant success [@REF].
These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.
However, ensemble forecasts have a range of potential downsides. The most obvious is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project. Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumption.
In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions. 
This means that non-parametric emulation is less suitable for use in epidemiological forecasting though for epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR].
In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources, and an easily generalised implementation.

@@

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
Real-time evaluation has shown that each of these classes of models may perform well though in general semi-mechanistic models, and models that include human insights have outperformed other approaches during certain periods [@REF].
We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF] with our forecasts performing relatively well compared to the majority of those submitted but rarely outperforming significantly.
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].
Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF]. 
Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts.
They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence.
On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts.
Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting.
There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.

Based on these insights from observing the performance of our forecast submissions and from the hub ensembles over time here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating, this means that the impact of daily reporting artefacts is mitigated.
It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model. 
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.
The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance.
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase. 
The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model.
To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence. 
The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion. 
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research.
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.

In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action. 
We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
Finally, we summarise our findings, discuss their implications, and highlight areas for potential study. We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models

@@

  We extracted forecasts and data on notified weekly COVID-19 cases from the ECDC forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
  We included all locations covered by the ECDC forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
  Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
  We used the latest available observed data as of the 20th of August 2022. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
  This follows the common practice of the forecasting hub projects in treating the latest available data as the forecast target regardless of the number of potential data revisions.
  Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

  The ECDC forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
  No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
  Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
  Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
  The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
  Submitted forecasts and target observations are available from the ECDC forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.

@@

\begin{align*}
  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
  r_t &= \left(\mathcal{H}[a](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
\end{align*}

@@

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the ECDC forecast hub  [@REF].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% credible intervals and by quantile. 
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage, convergence was not assessed, but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

To download and manipulate forecasts from the ECDC forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@REF].

@@

In our study period, incidence rates across ECDC nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility. Many nations, such as the UK saw large BA.1 waves in January that then resulted in declining incidence rates through February.
From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2. This wave was typically characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
The end of our study period was characterised by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates. Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
In addition, in general, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability. Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the ECDC forecasting hub (all ECDC nations and the UK making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the ECDC baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the ECDC forecasting hub project).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

### Forecast calibration {-}

>>>>

---
title: Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic. However, these forecasts require a coordinated team to produce and are effectively black boxes. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. 
  Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19. We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
  We found that this surrogate model performed visibly similarly to a subset of example locations though with increased uncertainty. Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse but that median performance was more comparable. The majority of the difference in performance was linked to a few forecasts from a period of high incidence. We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict. The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution whilst the tails of the surrogate forecast distribution were responsible for the large majority of the difference in median forecast performance.
  We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble. However more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements. We also conclude that our findings are largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
bibliography: library.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
---

@@

# Introduction {-}[a][b]

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@REF].
These approaches have drawn heavily on the experience of other domains, for example, climate modelling [@REF], where ensembles of both multiple models and from multiple teams have a long history of providing superior forecasts.

At the same time, ensemble forecasts have a range of potential downsides.
The most significant is their considerable cost to produce as they typically require contributions from multiple independent teams, the development of several models, and a centralised group to run the ensembling project.
Forecasts from these ensembles can be hard to interpret as it is difficult to link them to the underlying model assumptions.
Additional issues with multi-team collaborations include a lack of feedback to those contributing forecasts that would allow them to improve their forecast approaches, maintaining incentives for forecasters to continue to contribute, and difficulty improving the quality of the ensemble [@REF].
Each of these may impact the long-term quality of the resulting ensemble forecasts.

Here, we focus on addressing the issues of resource cost, and interpretability.
In climate forecasting, emulation approaches have been used to circumvent some of the resource requirement issues by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components [@REF].
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that non-parametric emulation is less suitable for use in epidemiological forecasting (while noting that epidemiological scenario analysis in-sample emulation using non-parametric methods may have a role [@REFLSHTMTBEMULATOR]). 
In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, and limited computational resources with an easily generalised implementation.

In this work, we propose and evaluate a surrogate forecast model approach which seeks to emulate the observed behaviour of recent COVID-19 forecast ensembles using a minimal set of easily communicated assumptions, limited computational resources with an easily generalised implementation.

@@

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].
Most submitted forecasts use available daily data, along with potentially other leading predictors, to produce daily forecasts that are then aggregated into weekly forecasts.
Many use a range of outlier processing techniques to reduce reporting artefacts which can bias forecasts [@REF].
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF], and human insight based forecast models that may also include elements of other methods [@REF].
Real-time evaluation has shown that each of these classes of models may perform well, although in general, semi-mechanistic models and models that include human insights have outperformed other approaches during certain periods [@REF].
We have contributed a range of forecasts to these forecast collaborations generally focussed on semi-mechanistic statistical methods [@REF], and human insight-based forecasts [@REF].

Our forecasts have performed relatively well compared to the majority of other submitted forecasts, but rarely outperforming significantly.
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@REF].
Based on our observations our forecasts have generally captured changes in trend better than most submitted models that do not use human insight based on known policy changes, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@REF]. 
Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts. They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence. On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts. Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting. There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models.

Based on the above insights from observing the performance of our forecast submissions and from the hub ensembles over time, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@REF] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

The second simplification we make is to ignore the underlying latent infection process and focus only on the observed reported cases.
This reduces computational requirements and removes the need for external information on the delay from infection to report which can potentially be misspecified leading to potentially reduced forecast performance. 
However, this sacrifices some of the interpretability of the forecast model as any epidemiological summary statistics we now calculate will be based on reported cases and not latent infections.
As discussed in [@gostic] this leads to varying amounts of bias depending on the epidemic phase.

The final simplification we make is to model the growth rate as a differenced auto-regressive process with order 1 rather than using a gaussian process-based method as we have done in other work [@EpiNow2].
This represents a parsimonious approach in that we encode our expectation that the growth rate should vary over time and allow this to influence the forecast but we include only a single lag term, reducing the computational overhead of the model. 
To model potential unobserved interventions and more general changes in transmission, we include an additional growth rate modifier restricted to be between 0 and 1 that differs depending on if the growth rate is positive or negative (due to potential differing responses when cases are growing or increasing) and that acts in a multiplicative fashion (meaning that larger absolute growth rates are reduced to zero growth more rapidly).
This reflects a simplified interpretation of how the ensemble appears to react to potential future changes by assuming a gradual return to stable incidence.

The only observation for which we do not make an adaptation is the apparent sharpness of the ensemble compared to our prior forecasting models choosing instead to make use of a negative binomial observation model allowing the inclusion of overdispersion.
This choice is motivated by our belief that the underlying transmission process is an exponential one and therefore a count error model (where variance is linked to the mean)  is a sensible choice though this is an area of current research. 
We suggest that part of the reason the hub ensembles exhibit this behaviour is due to the penalisation of overprediction compared to underprediction caused by the use of a generalised form of absolute error for the majority of forecast evaluations.

In this study, we evaluate an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions. 
We submitted this model to the European Forecast Hub and here we evaluate its performance relative to the Hub ensemble. 
First, we define the model and summarise its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow.  
We briefly outline some of the summary statistics it produces and their epidemiological relevance.  We then evaluate its real-time performance in comparison to the European forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed.  
Finally, we summarise our findings, discuss their implications, and highlight areas for potential study. We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models.

@@

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@REF].
We used the latest available observed data as of the 20th of August 2022. As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
This follows the practice of the European forecasting hub project in treating the latest available data as the forecast target regardless of the number of potential data revisions [@REF].
Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.

The European forecast hub requests forecasts for one to four-week forecast horizon and requires forecasts to use a pre-specified format with 23 quantiles of the predictive probability distribution.
No restrictions were placed on who could submit forecasts and the hub team actively invited participation from research groups known to be involved with COVID-19 forecasting projects.
Forecasters used a wide range of forecasting methods including statistical approaches, mechanistic models including deterministic ordinary differential equation models and agent-based models, semi-mechanistic models, expert elicitation models, and ensembles of multiple approaches [@https://covid19forecasthub.eu/community.html].
Teams submitted forecasts at the latest two days after the complete dataset for the forecast week became available and were allowed to use all data available at the time of submission (i.e including up to two days of data for the current week). 
The ensemble forecast was constructed by taking the median of all predictive quantiles without the exclusion of any validly submitted forecast (where validity was defined as passing minimal formatting checks by the hub team and timely submission). 
An ensemble was only produced for locations with at least 3 independent forecast models including the hub baseline model.
Submitted forecasts and target observations are available from the European forecast hub GitHub repository [@REF]. We provide code in the repository of this study to streamline access.

@@

\begin{align*}
  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
  r_t &= \left(\mathcal{H}(r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
\end{align*}

@@

## Forecast evaluation[c][d] {-}

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub  [@REF].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizo[e]n.

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile.[f] 
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmdstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage, convergence was not assessed, but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

To download and manipulate forecasts from the European forecasting hub [@EuroHub] we use  the `data.table` (`1.14.2`) [@data.table] and `gh` (`1.3.0`) [@gh] R packages.
We make use of further functionality from the `forecast.vocs` R package [@forecast.vocs] to prepare data for forecasting, visualise forecasts and summary measures, and summarise forecasts.
Forecast evaluation is implemented using the `scoringutils` R package (1.0.0) [@scoringutils], and the `scoringRules` R package (1.0.1) [@REF].

@@

In our study period, incidence rates across European nations and in the UK were primarily driven by the spread of novel subvariants of concern related to the Omicron variant and changes in population susceptibility. Many nations, such as the UK, saw large BA.1 waves in January that then resulted in declining incidence rates through February (Figure \@ref(fig:vis-forecasts)).
From late February, through to the end of May, most nations then saw another wave, typically with lower reported incidence rates, driven by BA.2. This wave was typically characterised by a lower peak than the BA.1 wave with a more gradual decrease in incidence.
The end of our study period was characterised by the gradual take-over of the BA.4/BA.5 subvariants that again had a lower peak and lower absolute growth rates. Unlike earlier periods in the pandemic, our study period did not see the use of new non-pharmaceutical interventions (NPIs) in response to increasing COVID-19 incidence in most locations.
In addition, ascertainment rates likely reduced over time in most locations due to reductions in routine testing, and reductions in test availability. Whilst both the reduced use of NPIs and testing generally occurred across nations our study period also marked an increase in the heterogeneity of the response to the COVID-19 pandemic with nations changing policy at different times and to different degrees.
This is in contrast to the early COVID-19 pandemic response for which most nations took similar actions at similar times.

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, `r forecast_dates` forecast dates from `r models` independent forecast models (including the European baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

### Forecast calibration[g][h] {-}
